Understanding the nanoscale structure of hexagonal phase lyotropic liquid crystal membranes

### 2018-06-11 Ben_Manuscripts/structure_paper/Draft.tex
comments.

  Nanostructured porous membranes made from the cross-linked hexagonal phase of
  self-assembled lyotropic liquid crystals (LLCs) are a promising material for
  selective separations.  In this work, we develop an atomistic molecular model
  of an LLC membrane. We show that our model is maximally consistent with
  experimental observations by comparing simulated X-ray Diffraction (XRD)
  patterns and calculated ionic conductivity to their experimental counterparts.
  We explore, in depth, the composition and structure of the nanopores in order
  to give insights that are not easily accessible to experimentalists.  The
  clearer picture of the nanoscopic structure of these membranes provided in this
  study will enable a better understanding of the mechanisms of small molecule
  transport within these nanopores.

  \end{abstract}

@@

  A highly selective membrane would be useful for the recovery of valuable
  products from complex aqueous and organic solutions.  For example, flowback
  water (FW) produced during hydraulic fracturing is a complex wastewater full of
  potentially valuable dissolved organic compounds such as acetate
  \cite{dischinger_application_2017}. There is increasing pressure to reuse
  hydraulic fracturing water rather than dispose of it in order to reduce social
  and environmental impacts as well as cost \cite{theodori_hydraulic_2014}.
  Rather fully than dispose of the wastestream generated in the recycling
  process, we can instead use highly selective membranes in order to successfully
  recover useful compounds.

@@

% MRS6: I would keep a paragraph about here how current membranes DON'T
% work for this purpose. But one paragraph is enough.
% BJC5: added back below paragraph
  Current commercial RO and NF membranes suffer limitations to there selectivity 
  which is inherent to their fabrication. Although scalable, each of these two types
  of membranes has a degree of stochasticity that makes overcoming well-known the 
  permeability-selectivity tradeoff a challenge. Namely, it is difficult to increase
  the permeability of a desired molecular or atomic species, while maintaining the same 
  retention of an undesired species \cite{werber_materials_2016}.

@@

  Crosslinked lyotropic liquid crystal (LLC) membranes may be capable of
  performing highly selective separations. LLCs are amphiphilic molecules that
  have the ability to self-assemble into porous nanostructures
  \cite{smith_ordered_1997} which can be crosslinked to create mechanically
  strong membrane films with pores on the order of 1 nm in diameter
  \cite{zhou_supported_2005}. Unlike most commercial NF membranes, LLC membrane pores
  are uniform in size. Since LLC membranes lack a pore size distribution, they
  inherently exhibit high selectivity due to their strict molecular weight
  cut-off (MWCO)~\cite{zhou_supported_2005}. Additionally, the LLC monomers
  described in this paper are salts, and therefore lead to Donnan exclusion of ions
  in solution. The membrane gains a net surface charge when counterions from the head
  groups that line the pore walls escape into the feed solution in order to balance 
  the gradients of concentration and electric potential \cite{donnan_theory_1995}.

@@

  % BJC: I like this figure, but I don't know if it is necessary in the main body. i  % Maybe move to supplemental. Thoughts?
  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
        \centering
        \includegraphics[trim=-2cm 0 0 0, clip, height=4.5cm]{Dibrpyr14.eps}
        \includegraphics[height=4.5cm]{bcc_phases.png}
        \caption{}\label{fig:bcc_phases}
  \end{subfigure}
  \hspace{1cm}
  \begin{subfigure}{0.45\textwidth}
        \centering
        \includegraphics[trim=0 -0.5cm 0 -0.5cm, clip, height=4.5cm]{NaGA3C11.eps}
        \includegraphics[height=4.5cm]{hexagonal_packing.png}
        \caption{}\label{fig:hII_phase}
  \end{subfigure}
%MRS5: not sure Q should be in here.
  \caption{The choice of LLC monomer and solvent leads to different accessible liquid
        crystalline phases. The monomer shown in (a) forms the Q\textsubscript{I} phase
        in either the Ia3d (left) or Pn3m (right) space group. The surfaces shown
        represent the interface between the hydrophilic and hydrophobic regions of the unit
        cell. As pictured, we assume that each region occupies an equal volume. The
        monomer shown in (b) will form the hexagonal columnar phase
        (H\textsubscript{II} in the presence of water or Col\textsubscript{h} when no
         water is present) The blue region represent the hydrophilic pores while the cyan region represents the surrounding hydrophobic monomer tails.}\label{fig:bcc_v_hII}
  \end{figure}

@@

  Our current understanding of the molecular details of LLC membrane nanostructure
  is not sufficient to be able to precisely design them for specific separations. 
  Over the past 20 years, H\textsubscript{II}-phase LLC polymer membrane studies have
  been limited primarily to Na-GA3C11 with some characterization done after minor 
  structural modifications. Resel et al.~varied the length of the monomer tails and the
  counterion used and observed its affect on pore spacing \cite{resel_structural_2000}.
  In a later study of rejection performance, it was shown that membranes formed by cross-
  linked Na-GA3C11 in the H\textsubscript{II} phase cannot separate solutes less than 1.2 
  nm in diameter because the pores are too large \cite{zhou_supported_2005}. We do not
  yet understand how to controllably reduce the effective pore size or how to tune
  the chemical environment in the nanopores of this or related materials for
  small molecule separations. The only source of predictive modeling for LLC
  systems have been macroscopic models that likely do not adequately describe
  transport at these length scales \cite{hatakeyama_water_2011}. It will be
  challenging to efficiently narrow down the large design space in a laboratory
  setting without a robust model.

  A molecular-level understanding of LLC polymer membrane structure, enabled by
  molecular dynamics (MD) simulations, can provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. A good molecular model should incorporate a detailed picture of the
  nanoscopic pore structure which will be crucial to understanding the role of
  monomer structure in solute transport and membrane design. Models resulting
  from molecular dynamics simulations will provide the required level of detail
  (Fig. \ref{fig:detail}), assuming the force fields are sufficiently accurate.
  With such an atomistic model, we can directly observe molecular-level solute
  transport and suggest governing mechanisms. We can observe how the choice of
  head group interacts with solutes of interest. We can interchange
  counterions which may influence both the pore size and the strength of the
  Donnan potential. 
  %which affects the degree to which the membrane can exclude
  %charged species at the membrane surface.

@@

  \begin{figure}[!htb]
  \centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{cartoon_pore.png}
		\caption{}~\label{fig:undetailed_pore}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{detailed_pore.png}
		\caption{}~\label{fig:detailed_pore}
	\end{subfigure}
    \caption{One can only speculate about solute behavior inside the membrane
	  pores based on the previously established picture of the pore structure (a). 
      We use a detailed molecular model in order to appropriately model the pore's
      complex architecture which is crucial to understanding the mechanism of solute
	  transport (b). The head group region is colored blue and the tail region is
	  colored cyan in both representations. This image was generated from the molecular
	  simulations run for this paper.}~\label{fig:detail}
  \end{figure}

  Here we develop a molecular model of the Col\textsubscript{h} assembly formed by
  Na-GA3C11. Compared to the H\textsubscript{II} phase, the Col\textsubscript{h}
  phase is a simpler starting point. The system has no water which will allow us
  to simulate longer timescales, and there exists detailed experimental
  characterization of the fully aligned state, including 2D wide-angle X-ray
  scattering (WAXS) patterns (Figure\ref{fig:WAXS}) which are useful for reconstructing 
  structural data. We will use this Col\textsubscript{h} model in order to build an 
  H\textsubscript{II} model in the future.
  %The two phases appear to share similar structural characteristics since the
  %pore spacings in each system are in close agreement
  %\cite{feng_thin_2016,resel_structural_2000}.

@@

  \begin{figure}[!htb]
        \centering
        \begin{subfigure}[t]{0.43\linewidth}
                \centering
                \includegraphics[width=\linewidth]{SAXS.png}
                \caption{}\label{fig:SAXS}
        \end{subfigure}
        \begin{subfigure}[t]{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{WAXS_annotated_words.png} 
                \caption{}\label{fig:WAXS}
        \end{subfigure}
	\caption{(a) (Reproduced from \cite{feng_thin_2016}) The repeat spacing
		in the 1D small angle X-ray scattering pattern is characteristic of hexagonal
		packing. The leading peak, q*, represents the distance between the
		d\textsubscript{100} planes. Using this distance, we know that the distance
		between pore centers is 4.12 nm. (b) 2D WAXS gives
		details about repeating features on the order of angstroms. Experimentalists
		have explained each of the 5 major reflections present as follows: (R-$\pi$) Aromatic
		head groups $\pi-\pi$ stack 3.7 \AA~apart. (R-double) Monomers arrange vertically in
		a 2\textsubscript{1} helix. (R-alkanes) Alkane chain tails pack 4.5 \AA~apart. (R-spots)
		Monomer tails are tilted with respect to the membrane plane. (V) As derived from
		SAXS, the pores are spaced 4.12 nm apart and pack hexagonally}
	\label{fig:SWAXS}
 \end{figure}

@@

 	We are careful to avoid the term 'layers' since a liquid crystalline system has, by 
 	definition, long range order in 1 or 2 spatial dimensions and short range order in 	
 	the other dimensions.~\cite{chaikin_principles_1995}  % page 58
 	In the system we are studying, there are long-range 2D correlations in the 
 	hexagonal array of pores (xy plane) and short range correlations in the z-direction.
 	We will use our atomistic molecular model to study how the system's structure is 
 	effected by the density of monomers surrounding each pore's hydrophilic core.

@@

	On the short timescales accessible to MD, we observe distinct metastable configurations
	which depend on starting configuration. We simulated XRD patterns of our system and 
	compared them to experimental 2D WAXS patterns (Figure~\ref{fig:WAXS} so that we ensure our
	model creates a nanoscopic chemical environment maximally consistent with experiment within 
	the constraints	of our forcefield. Using this approach, we are able to confirm some previous
	interpretations	of the WAXS pattern and refute others.

@@

  We parameterized the liquid crystal monomer Na-GA3C11 using the Generalized AMBER
  Force Field (GAFF) \cite{wang_development_2004} with the Antechamber package
  \cite{wang_automatic_2006} provided with AmberTools16
  \cite{case_ambertools16_2016}. We assigned atomic charges using the am1bccsym
  method of \texttt{molcharge} shipped with QUACPAC from Openeye Scientific
  Software. We ran all molecular dynamics simulations using Gromacs 2016
  \cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}.

@@

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended
  periodically. Each pore is made of columns of stacked monomers with periodic
  continuity along the pore axis, avoiding any edge effects and creating an
  infinite length pore ideal for studying transport. We prefer a small number stacked
  monomers in order to reduce computational cost and to allow us to look at
  longer timescales. Ultimately, we chose to build a system with 20 monomers
  per column in order to obtain sufficient resolution when simulating
  X-ray diffraction patterns. %This point will be explained in more detail later.

@@

  We chose the vertical spacing between monomers for the initial configuration based
  on R-$\pi$ and then allowed the system to readjust during equilibration. Each monomer was
  rotated so the plane of its aromatic head groups would be coplanar with the xy plane. We
  explored three different initial monomer spacings. The first is exactly
  equal to R-$\pi$ with layers placed so aromatic rings stack 3.7 \AA~apart in
  the z-direction. We explore a second system with an initial spacing of 5
  \AA. We briefly explored a third system with an initial spacing of 10
  \AA, however it shows non-physical behavior which is detailed in the 
  Supporting Information, section S-\ref{S-section:initial_dbwl}.

@@

  \subsection{Equilibration}

  We developed equilibration schemes to create dry and wet configurations. Both
  schemes start with an initial configuration generated according to the previous
  guidelines. To create a dry configuration, we fix monomer head groups in the
  sandwiched or parallel-displaced configuration using position restraints with a
  force constant of 10$^6$ kJ mol$^{-1}$ nm$^{-2}$. We run a 50 ps simulation in
  the NVT ensemble which allows the monomer tails to settle without disrupting
  the ordering of the head groups. Doing so also mitigates system dependence on
  initial monomer configuration. Every 50 ps, we reduce the force constants by
  the square root of its previous value. Once the force constant is below 10 KJ
  mol$^{-1}$ nm$^{-2}$, we reduce the restraints in a sequence with values of
  8, 3, 2, 1, and 0 KJ mol$^{-1}$ nm$^{-2}$ respectively. We allow the resulting
  unrestrained structure to equilibrate for 5 ns in the NPT ensemble
  with pressure controlled by the berendsen barostat. Next, we run long NPT
  equilibration simulations for at least 400 ns using the Parrinello-Rahman
  barostat with a time constant of 10 ps.

@@

  Using equilibrated structures, we carry out various calculations to
  characterize the system. We define the point at which a system is equilibrated
  based on when the distance between pores stops changing.  We determined when
  the distances stopped changing by applying the statistical test,
  \texttt{pymbar.timeseries.detectEquilibration}, to the time series
  \cite{chodera_simple_2016,shirts_statistically_2008}. Simulations of 400 ns
  give at least 50 ns of equilibrated simulation trajectory.

@@

  We measured the one dimensional pair distribution function, $g(z)$, between centers 
  of masses of aromatic head group rings along the z-axis (perpendicular to
  the membrane plane). We averaged all 1D slices in the z-direction of the full 3D 
  correlation function within 2.1~\AA~of $(x, y)=(0, 0)$. We chose 2.1 \AA as a crude 
  approximation of the radius of the phenyl ring plane. 
  %In this way, we include any instances of 
  %overlapping electron density which would contribute to R-$\pi$. 
  We calculated the radius as the sum of the longest C-C distance within a phenyl 
  ring (2.8~\AA) and two times the carbon atom radius (0.7~\AA).

@@

  We also used $g(z)$ to calculate the equilibrated vertical stacking distance between
  monomers, $\mathit{d}_equil$. 
  %BJC5: in case there is an criticism of this method
  %The simplest way to do so, would be to fourier transform
  %$g(z)$ and extract the highest amplitude frequency of the resulting fourier series. 
  %However, the size of our system limits the width of the bins in fourier space. The bins
  %are too large to extract a precise value of $\mathit{d}_equil$. Instead, 
  We fit a decaying sinusoidal function to $g(z)$ of the form:
  \begin{equation}
  	1 - Acos(\frac{2\pi}{\mathit{d}_{equil}}z + B)e^{-z/L}
  	\label{eqn:decaying_sinusoid}
  \end{equation}
  where A and B are fit parameters for the function's amplitude and phase shift respectively.
  This function could be used in place of Equation~\ref{eqn:decaying_exponential}, however
  it does not fit the peaks of $g(z)$ from parallel displaced configurations well enough
  to extract a reliable value of L.
  %BJC5: maybe include plots of these fits in supporting info

@@

  In order to fully match synthetic procedures, we created a cross-linking
  algorithm that one can apply to equilibrated structures. The purpose of
  cross-linking is to maintain macroscopic alignment of the crystalline domains,
  ensuring aligned, hexagonally packed pores. For that reason, we are not
  concerned with replicating the kinetics of the reaction, but instead emphasize
  the consistency of the final structure with experimental structural data.

  %BJC5: perhaps this paragraph should be moved to supporting info.
  We developed the algorithm based on the known reaction mechanism.
  Cross-linking of this system is a free radical polymerization (FRP) taking
  place between terminal vinyl groups present on each of the three monomer tails.
  The reaction mechanism is drawn out in Figure S-\ref{S-fig:xlink_mech}.
  FRPs require an initiator which bonds to the system, meaning new atoms are
  introduced into the system. For simplicity, we simulated the initiator as
  hydrogen and made it present in the simulation by including them as dummy atoms
  in all possible locations where an addition could occur. We carry out the
  cross-linking procedure iteratively. During each iteration, the algorithm
  selects eligible bonding carbon atoms based on a distance cut-off. The topology
  is updated with new bonds and dummy hydrogen atoms are changed to appropriate
  hydrogen types.  Head-to-tail addition was the only propagation mode considered
  due to its dominance in the real system \cite{young_introduction_2011}. We did
  not consider direction of attack because the resultant mixture is racemic.

@@

  %BJC5: thesis needs rewording.
  Disordered basin systems built with 6 columns-per-pore are not compact in the z-direction.
  The final pore spacing is close to experiment within its uncertainty. Relative to ordered basin 
  6 column-per-pore systems, the unit cell is elongated in its z dimension and contracted in its
  x and y dimensions, however the box volume stays nearly constant in all cases 
  (Figure~\ref{fig:box_dimensions}). The system is likely not fully equilibrated and may eventually
  rearrange into something that resembles a 5 column-per-pore system. Further investigations 
  into this are beyond the scope of this paper.

>>>>

  Nanostructured porous membranes made from the cross-linked hexagonal phase of
  self-assembled lyotropic liquid crystals (LLCs) are a promising material for
  selective separations.  In this work, we develop an atomistic molecular model
  of an LLC membrane. We show that our model is maximally consistent with
  experimental observations by comparing simulated X-ray Diffraction (XRD)
  patterns and calculated ionic conductivity to their experimental counterparts.
  We explore, in depth, the composition and structure of the nanopores in order
  to give insights that are not easily accessible to experimentalists.  The
  clearer picture of the nanoscopic structure of these membranes provided in this
  study will enable a better understanding of the mechanisms of small molecule
  transport within these nanopores.
  %MRS7: some more of the conclusions should be in the abstract.  'introduction' just raises questions, but 
  %'abstract' should give a executive-level summary of the answers.
  \end{abstract}

@@

  More highly selective membranes would be extremely useful for the recovery of valuable
  products from complex aqueous and organic solutions.  For example, flowback
  water (FW) produced during hydraulic fracturing is a complex wastewater full of
  potentially valuable dissolved organic compounds such as acetate
  \cite{dischinger_application_2017}. There is increasing pressure to reuse
  hydraulic fracturing water rather than dispose of it in order to reduce social
  and environmental impacts as well as cost \cite{theodori_hydraulic_2014}.
  Rather fully than dispose of the wastestream generated in the recycling
  process, we can instead use highly selective membranes in order to successfully
  recover useful compounds.

  %MRS: abstract seems a bit limited, describe more generally (and
  briefly!) other things selective membranes are good for. The paper is primarily about selective membranes, not fracking water.  It seems out of balance to be talking about only one application.

@@

% MRS6: I would keep a paragraph about here how current membranes DON'T
% work for this purpose. But one paragraph is enough.
% BJC5: added back below paragraph.  
% MRS7: Do a pass for clarity at the sentence level.  For example ``degree of stochasticity'' is vague.
  Current commercial RO and NF membranes suffer limitations to their selectivity 
  which is inherent to their fabrication. Although scalable, each of these two types
  of membranes has a degree of stochasticity that makes overcoming well-known the 
  permeability-selectivity tradeoff a challenge. Namely, it is difficult to increase
  the permeability of a desired molecular or atomic species, while maintaining the same 
  retention of an undesired species \cite{werber_materials_2016}.

@@

  Crosslinked lyotropic liquid crystal (LLC) membranes may be capable of
  performing highly selective separations. LLCs are amphiphilic molecules that
  have the ability to self-assemble into porous nanostructures
  \cite{smith_ordered_1997} which can be crosslinked to create mechanically
  strong membrane films with pores on the order of 1 nm in diameter
  \cite{zhou_supported_2005}. Unlike most commercial NF membranes, LLC membrane pores
  are uniform in size because they are self-assembled. 
  Since LLC membranes lack a pore size distribution, they
  inherently exhibit high selectivity due to their strict molecular weight
  cut-off (MWCO)~\cite{zhou_supported_2005}. Additionally, the LLC monomers
  described in this paper are salts, and therefore lead to Donnan exclusion of ions
  in solution. The membrane gains a net surface charge when counterions from the head
  groups that line the pore walls escape into the feed solution in order to balance 
  the gradients of concentration and electric potential \cite{donnan_theory_1995}.

@@

  % BJC: I like this figure, but I don't know if it is necessary in the main body. i  % Maybe move to supplemental. Thoughts?
  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
        \centering
        \includegraphics[trim=-2cm 0 0 0, clip, height=4.5cm]{Dibrpyr14.eps}
        \includegraphics[height=4.5cm]{bcc_phases.png}
        \caption{}\label{fig:bcc_phases}
  \end{subfigure}
  \hspace{1cm}
  \begin{subfigure}{0.45\textwidth}
        \centering
        \includegraphics[trim=0 -0.5cm 0 -0.5cm, clip, height=4.5cm]{NaGA3C11.eps}
        \includegraphics[height=4.5cm]{hexagonal_packing.png}
        \caption{}\label{fig:hII_phase}
  \end{subfigure}
%MRS5: not sure Q should be in here.  Could cut length by removing some of the detail about Q phase, since it 
isn't discussed in the paper. 
  \caption{The choice of LLC monomer and solvent leads to different accessible liquid
        crystalline phases. The monomer shown in (a) forms the Q\textsubscript{I} phase
        in either the Ia3d (left) or Pn3m (right) space group. The surfaces shown
        represent the interface between the hydrophilic and hydrophobic regions of the unit
        cell. 
        %MRS7: not needed
        %As pictured, we assume that each region occupies an equal volume. 
        The
        monomer shown in (b) will form the hexagonal columnar phase
        (H\textsubscript{II} in the presence of water or Col\textsubscript{h} when no
         water is present) The blue region represent the hydrophilic pores while the cyan region represents the surrounding hydrophobic monomer tails.}\label{fig:bcc_v_hII}
  \end{figure}

@@

  Our current understanding of the molecular details of LLC membrane nanostructure
  is not sufficient to be able to precisely design them for specific separations. 
  Over the past 20 years, H\textsubscript{II}-phase LLC polymer membrane studies have
  been limited primarily to Na-GA3C11 with some characterization done after minor 
  structural modifications. Resel et al.~varied the length of the monomer tails and the
  counterion used and observed its affect on pore spacing \cite{resel_structural_2000}.
  In a later study of rejection performance, it was shown that membranes formed by cross-
  linked Na-GA3C11 in the H\textsubscript{II} phase cannot separate solutes less than 1.2 
  nm in diameter because the pores are too large \cite{zhou_supported_2005}. We do not
  yet understand how to controllably reduce the effective pore size or how to tune
  the chemical environment in the nanopores of this or related materials for
  small molecule separations. The only source of predictive modeling for LLC
  systems have been macroscopic models that likely do not adequately describe
  transport at these length scales \cite{hatakeyama_water_2011}. 
  %MRS7: maybe a pass at this?
  Modeling with molecular detail could provide sufficient information about the mechanisms 
  and chemical features to better inform experimental design of similar nanostructured membranes. 
  %It will be
  %challenging to efficiently narrow down the large design space in a laboratory
  %setting without a robust model.

  %MRS7: changing future tense to present tense here.
  A molecular-level understanding of LLC polymer membrane structure, enabled by
  molecular dynamics (MD) simulations, can provide guidelines to reduce the large
  chemical space available to design monomers for creation of separation-specific
  membranes. A good molecular model should incorporate a detailed picture of the
  nanoscopic pore structure which is crucial to understanding the role of
  monomer structure in solute transport and membrane design. Models resulting
  from molecular dynamics simulations can provide the required level of detail
  (Fig. \ref{fig:detail}), assuming the force fields are sufficiently accurate.
  With such an atomistic model, we can directly observe molecular-level solute
  transport and suggest governing mechanisms. We can observe how the choice of
  head group interacts with solutes of interest. We can interchange
  counterions which may influence both the pore size and the strength of the
  Donnan potential. 
  %which affects the degree to which the membrane can exclude
  %charged species at the membrane surface.

@@

  \begin{figure}[!htb]
  \centering
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{cartoon_pore.png}
		\caption{}~\label{fig:undetailed_pore}
	\end{subfigure}
	\begin{subfigure}{0.45\linewidth}
		\centering
		\includegraphics[width=\textwidth]{detailed_pore.png}
		\caption{}~\label{fig:detailed_pore}
	\end{subfigure}
    %MRS7: Adjusted the phrasing
    \caption{(a) Previous understanding of the pores are essentially speculations based on limited chemical and experimental data. (b) We use a detailed molecular model in this paper in order to appropriately model the pore's
      complex architecture which is crucial to understanding the mechanism of solute
	  transport. In both pictures, the head group region is colored blue and the tail region is
	  colored cyan. 
          %MRS7: don't think this is needed.
          %This image was generated from the molecular simulations run for this paper.
          }~\label{fig:detail}
  \end{figure}

  %Here 
  In this paper,
   we develop a molecular model of the Col\textsubscript{h} assembly formed by
  Na-GA3C11. Compared to the H\textsubscript{II} phase, the Col\textsubscript{h}
  phase is a simpler starting point. The system has no water which will allow us
  to simulate longer timescales, and there exists detailed experimental
  characterization of the fully aligned state, including 2D wide-angle X-ray
  scattering (WAXS) patterns (Figure~\ref{fig:WAXS}) which are useful for reconstructing 
  structural data. We will use this Col\textsubscript{h} model in order to build an 
  H\textsubscript{II} model in the future.
  %The two phases appear to share similar structural characteristics since the
  %pore spacings in each system are in close agreement
  %\cite{feng_thin_2016,resel_structural_2000}.

@@

  \begin{figure}[!htb]
        \centering
        \begin{subfigure}[t]{0.43\linewidth}
                \centering
                \includegraphics[width=\linewidth]{SAXS.png}
                \caption{}\label{fig:SAXS}
        \end{subfigure}
        \begin{subfigure}[t]{0.47\linewidth}
                \centering
                \includegraphics[width=\linewidth]{WAXS_annotated_words.png} 
                \caption{}\label{fig:WAXS}
        \end{subfigure}
        %MRS7: picture says R-helix still.
	\caption{(a) (Reproduced from \citenum{feng_thin_2016}) The repeat spacing
		in the 1D small angle X-ray scattering pattern is characteristic of hexagonal
		packing. The leading peak, q*, represents the distance between the
		d\textsubscript{100} planes. Using this distance, we know that the distance
		between pore centers is 4.12 nm. (b) 2D WAXS gives
		details about repeating features on the order of angstroms. Experimentalists
		have explained each of the 5 major reflections present as follows: (R-$\pi$) Aromatic
		head groups $\pi-\pi$ stack 3.7 \AA~apart. (R-double) Monomers arrange vertically in
		a 2\textsubscript{1} helix. (R-alkanes) Alkane chain tails pack 4.5 \AA~apart. (R-spots)
		Monomer tails are tilted with respect to the membrane plane. (V) As derived from
		SAXS, the pores are spaced 4.12 nm apart and pack hexagonally}
	\label{fig:SWAXS}
 \end{figure}

@@

        %MRS7: is this really a liquid crystalline system, even if it is made of liquid crystal monomers? Should we say
        % in many liquid crystal systems?
 	We are careful to avoid the term 'layers' since a liquid crystalline system has, by 
 	definition, long range order in 1 or 2 spatial dimensions and short range order in 	
 	the other dimensions.~\cite{chaikin_principles_1995}  % page 58
 	In the system we are studying, there are long-range 2D correlations in the 
 	hexagonal array of pores (xy plane) and short range correlations in the z-direction.
        %MRS7: look carefully at tenses.  Don't describe the research in future tense. 
        % if it's describing your presentation of the research in the paper , use present tense. 
        % If it's describing something you did, use past tense. I changed some of them, but not all.
 	%We will use 
        In this study, we use
        our atomistic molecular model to study how the system's structure is 
        affected
 	%effected 
        by the density of monomers surrounding each pore's hydrophilic core.

@@

	On the short timescales accessible to MD
        %MRS7: added the following to make it clear the scale of the effort
        (even the 100's of nanoseconds of simulation peformed here are short compared to experimental timescales),
        we observe distinct metastable configurations
	which depend on starting configuration. We simulated XRD patterns of our system and 
	compared them to experimental 2D WAXS patterns (Figure~\ref{fig:WAXS}) so that we ensure our
	model creates a nanoscopic chemical environment maximally consistent with experiment within 
	the constraints	of our forcefield. Using this approach, we are able to confirm some previous
	interpretations	of the WAXS pattern and refute others.

@@

  We parameterized the liquid crystal monomer Na-GA3C11 using the Generalized AMBER
  Force Field (GAFF) \cite{wang_development_2004} with the Antechamber package
  \cite{wang_automatic_2006} provided with AmberTools16
  \cite{case_ambertools16_2016}. We assigned atomic charges using the am1bccsym
  method of \texttt{molcharge} shipped with QUACPAC from Openeye Scientific
  Software. We ran all molecular dynamics simulations using GROMACS 2016
  \cite{bekker_gromacs:_1993,berendsen_gromacs:_1995,van_der_spoel_gromacs:_2005,hess_gromacs_2008}.

@@

  A typical simulation volume contains four pores in a monoclinic unit cell,
  the smallest unit cell that maintains hexagonal symmetry when extended
  periodically. Each pore is made of columns of stacked monomers with periodic
  continuity along the pore axis, avoiding any edge effects and creating an
  infinite length pore ideal for studying transport. We prefer a small number stacked
  monomers in order to reduce computational cost and to allow us to look at
  longer timescales. Ultimately, we chose to build a system with 20 monomers
  per column in order to obtain sufficient resolution when simulating
  X-ray diffraction patterns. %This point will be explained in more detail later.
  %MRS7: refer to supporting information (if it's there) for more discussion on stability as a function of height?

@@

  We chose the vertical spacing between monomers for the initial configuration based
  on R-$\pi$ and then allowed the system to readjust during equilibration. Each monomer was
  rotated so the plane of its aromatic head groups would be coplanar with the xy plane. We
  explored three different initial monomer spacings. The first is exactly
  equal to R-$\pi$ with layers placed so aromatic rings stack 3.7 \AA~apart in
  the z-direction. We explore a second system with an initial spacing of 5
  \AA. We briefly explored a third system with an initial spacing of 10
  \AA. However this spacing yields non-physical behavior which is detailed in the 
  Supporting Information, section S-\ref{S-section:initial_dbwl}.

@@

  \subsection{Equilibration}
  %MRS7: could put this mostly in supporting. 
  We developed equilibration schemes to create dry and wet configurations. Both
  schemes start with an initial configuration generated according to the previous
  guidelines. To create a dry configuration, we fix monomer head groups in the
  sandwiched or parallel-displaced configuration using position restraints with a
  force constant of 10$^6$ kJ mol$^{-1}$ nm$^{-2}$. We run a 50 ps simulation in
  the NVT ensemble which allows the monomer tails to settle without disrupting
  the ordering of the head groups. Doing so also mitigates system dependence on
  initial monomer configuration. Every 50 ps, we reduce the force constants by
  the square root of its previous value. Once the force constant is below 10 KJ
  mol$^{-1}$ nm$^{-2}$, we reduce the restraints in a sequence with values of
  8, 3, 2, 1, and 0 KJ mol$^{-1}$ nm$^{-2}$ respectively. We allow the resulting
  unrestrained structure to equilibrate for 5 ns in the NPT ensemble
  with pressure controlled by the berendsen barostat. Next, we run long NPT
  equilibration simulations for at least 400 ns using the Parrinello-Rahman
  barostat with a time constant of 10 ps.

@@

  Using equilibrated structures, we carry out various calculations to
  characterize the system. We define the point at which a system is equilibrated
  based on when the distance between pores stops changing.  We determined when
  the distances stopped changing by applying the statistical test,
  \texttt{pymbar.timeseries.detectEquilibration}, to the time series
  \cite{chodera_simple_2016,shirts_statistically_2008}. Simulations of 400 ns
  %MRS7: is there a typical total time as well?
  give at least 50 ns of equilibrated simulation trajectory.

@@

  We measured the one dimensional pair distribution function, $g(z)$, between centers 
  of masses of aromatic head group rings along the z-axis (perpendicular to
  the membrane plane). We averaged all 1D slices in the z-direction of the full 3D 
  correlation function within 2.1~\AA~of $(x, y)=(0, 0)$. We chose 2.1 \AA~as a crude 
  approximation of the radius of the phenyl ring plane. 
  %In this way, we include any instances of 
  %overlapping electron density which would contribute to R-$\pi$. 
  We calculated the radius as the sum of the longest C-C distance within a phenyl 
  ring (2.8~\AA) and two times the carbon atom radius (0.7~\AA).

@@

  We also used $g(z)$ to calculate the equilibrated vertical stacking distance between
  monomers, $d_{equil}$. 
  %BJC5: in case there is an criticism of this method
  %The simplest way to do so, would be to fourier transform
  %$g(z)$ and extract the highest amplitude frequency of the resulting fourier series. 
  %However, the size of our system limits the width of the bins in fourier space. The bins
  %are too large to extract a precise value of $\mathit{d}_equil$. Instead, 
  We fit a decaying sinusoidal function to $g(z)$ of the form:
  \begin{equation}
  	1 - A\cos\left(\frac{2\pi}{\mathit{d}_{equil}}z + B\right)e^{-z/L}
  	\label{eqn:decaying_sinusoid}
  \end{equation}
  where A and B are fit parameters for the function's amplitude and phase shift respectively.
  This function could be used in place of Equation~\ref{eqn:decaying_exponential}, however
  it does not fit the peaks of $g(z)$ from parallel displaced configurations well enough
  to extract a reliable value of L.
  %BJC5: maybe include plots of these fits in supporting info

@@

  %In order to fully match synthetic procedures, 
  In order to maximally match the experimental membrane synthesis process
  we created a cross-linking
  algorithm that one can apply to equilibrated structures. 
  %MRS7: I think the main purpose of crosslinking is mechanical stability.
  The purpose of
  cross-linking is to maintain macroscopic alignment of the crystalline domains,
  ensuring aligned, hexagonally packed pores. For that reason, we are not
  concerned with replicating the kinetics of the reaction, but instead emphasize
  %MRS7: Since the initial structure isn't that well captured by experiment, then it's probably
  % better to say that we are interested in knowing how much and in why way cross-linking modulates the structure. 
  the consistency of the final structure with experimental structural data.

  %BJC5: perhaps this paragraph should be moved to supporting info.
  %MRS7: probably?  Describing the _idea_ of the procedure can be in
  %main paper, but perhaps more details of the procedure in the supporting info.

  We developed the algorithm based on the known reaction mechanism.
  Cross-linking of this system is a free radical polymerization (FRP) taking
  place between terminal vinyl groups present on each of the three monomer tails.
  The reaction mechanism is drawn out in Figure S-\ref{S-fig:xlink_mech}.
  FRPs require an initiator which bonds to the system, meaning new atoms are
  introduced into the system. For simplicity, we simulated the initiator as
  hydrogen and made it present in the simulation by including them as dummy atoms
  in all possible locations where an addition could occur. We carry out the
  cross-linking procedure iteratively. During each iteration, the algorithm
  selects eligible bonding carbon atoms based on a distance cut-off. The topology
  is updated with new bonds and dummy hydrogen atoms are changed to appropriate
  hydrogen types.  Head-to-tail addition was the only propagation mode considered
  due to its dominance in the real system \cite{young_introduction_2011}. We did
  not consider direction of attack because the resultant mixture is racemic.

@@

  %BJC5: thesis needs rewording.
  %MRS7: probably, since the disorered - 6 are close to the experimental distances.  Is it accurate to say
  %that they don't have the experimental structural features? 
  Disordered basin systems built with 6 columns-per-pore are not compact in the z-direction.
  The final pore spacing is close to experiment within its uncertainty. Relative to ordered basin 
  6 column-per-pore systems, the unit cell is elongated in its z dimension and contracted in its
  x and y dimensions, however the box volume stays nearly constant in all cases 
  (Figure~\ref{fig:box_dimensions}). The system is likely not fully equilibrated and may eventually
  rearrange into something that resembles a 5 column-per-pore system. Further investigations 
  into this are beyond the scope of this paper.
